---
title: "Plant Stress Physiology: Drought FTSW, Salinity Ions, and Heat Shock Kinetics"
author: "Paul Lopez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plant Stress Physiology: Drought FTSW, Salinity Ions, and Heat Shock Kinetics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

Abiotic stress phenomena (drought, salinity, heat shock during anthesis Zadoks 65) exhibit non-linear threshold responses:

1. **Drought**: Fraction of Transpirable Soil Water (**FTSW**) stomatal closure breakpoint $\theta$.
2. **Salinity**: Ion homeostasis ($K^+/Na^+$ selectivity) and active osmotic adjustment ($\Delta \Psi_s^{100}$).
3. **Heat at Antesis**: Pollen sterility logit GLM and Richards grain-filling kinetics.
4. **Photosynthesis**: Farquhar-von Caemmerer-Berry ($A/C_i$) $V_{c\max}$ and $J_{\max}$.

---

```{r setup}
library(agriinsight)
set.seed(2026)
```

# 1. Drought FTSW Breakpoint Modeling

```{r ftsw, fig.width = 6.5, fig.height = 4.5}
ftsw_seq <- seq(1.0, 0.05, by = -0.05)
ntr_seq  <- ifelse(ftsw_seq >= 0.35, 1.0, 1.0 + 3.0 * (ftsw_seq - 0.35)) + rnorm(length(ftsw_seq), 0, 0.02)

ftsw_fit <- get_ftsw_breakpoint(ftsw_seq, ntr_seq)
ftsw_fit

if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot(ftsw_fit)
}
```

---

# 2. Salinity Ion Homeostasis & Osmotic Adjustment

```{r salinity}
# K+/Na+ homeostasis in durum wheat leaf blade
ions <- get_ion_homeostasis(k_tissue = 4800, na_tissue = 750, k_substrate = 1.0, na_substrate = 1.0)
ions

# Active osmotic adjustment standardized to 100% RWC
oa <- get_osmotic_adjustment(psi_s_stress = -2.2, rwc_stress = 0.65, psi_s_control = -1.1, rwc_control = 0.95)
oa
```

---

# 3. Photosynthetic $A/C_i$ Gas Exchange Curve

```{r aci}
ci_vals <- seq(50, 850, length.out = 10)
anet_vals <- c(4.8, 11.5, 17.8, 23.5, 27.8, 30.9, 32.8, 34.0, 34.8, 35.0)

params_fvcb <- get_photosynthetic_params(ci = ci_vals, a_net = anet_vals)
params_fvcb
```
