---
title: "Getting Started with agriinsight: Unified Biometrics for Agriculture & Biology"
author: "Paul Lopez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with agriinsight: Unified Biometrics for Agriculture & Biology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction

Statistical analysis in agricultural sciences and experimental biology spans diverse biological levels: from controlled environment chambers (CEA) and microclimatic gradients to multi-environment field breeding trials ($G \times E$), non-linear plant stress kinetics, and soil fertility calibration.

Existing R packages like `easystats/insight` provide model-agnostic parameter extractors for general linear models, but lack domain-specific biometric extractors required by agronomists and plant biologists (such as **Cullis generalized heritability**, **2D anisotropic semivariograms**, **FTSW stomatal breakpoints**, **AMMI/GGE stability metrics**, and **Piepho maximal clique CLD groupings**).

**`agriinsight`** bridges this gap by offering a zero-heavy-dependency S3 meta-interface that extracts biometrical parameters across `lme4`, `nlme`, `glmmTMB`, `SpATS`, `sommer`, and `drc`, accompanied by publication-grade visualizers and AI-ready reporting prompts.

---

# Installation & Loading

```r
# Install development version from GitHub
remotes::install_github("PALP31/agriinsight")
```

```{r setup}
library(agriinsight)
library(lme4)
```

---

# Built-in Realistic Datasets (`simulate_agri_data`)

`agriinsight` includes a built-in simulation suite reproducing real agricultural structures:

```{r datasets}
# 1. Greenhouse wheat trial with pad-to-fan thermal gradient
df_gh <- simulate_agri_data("greenhouse_wheat", n = 96)
head(df_gh, 4)

# 2. Multi-Environment Trial (MET) with GxE interactions
df_met <- simulate_agri_data("met_field_trials")
head(df_met, 4)
```

---

# Core Workflow: Controlled Environment Agriculture (CEA)

In greenhouse trials, physical cooling pads create microclimatic VPD and temperature gradients, while uneven germination creates density imbalances. We fit a linear mixed model accounting for these covariates and extract physical strata variances and Cullis heritability:

```{r greenhouse-fit}
# Fit LMM
mod_gh <- lmer(Biomass_g ~ Inoculant + Pad_Distance + Pot_Density + (1 | Genotype) + (1 | Table_Block), data = df_gh)

# Check microclimate gradient
grads <- get_greenhouse_gradients(mod_gh, y_axis = "Pad_Distance")
cat("Longitudinal gradient detected:", grads$gradient_detected, "\n")

# Extract variance by physical stratum
get_strata_variance(mod_gh)

# Calculate Cullis generalized heritability (H2)
get_heritability(mod_gh, type = "cullis")
```

---

# Compact Letter Displays (CLD) via Piepho (2004)

`agriinsight` implements the exact maximal clique algorithm (Bron-Kerbosch) in pure Base R without external dependencies:

```{r cld}
# Post-hoc comparisons with minimal letter codes
cld_tab <- agro_cld(mod_gh, term = "Inoculant", method = "tukey")
cld_tab
```
