---
title: "AI Diagnostic Co-Authoring & Publication-Ready Visual Exports"
author: "Paul Lopez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AI Diagnostic Co-Authoring & Publication-Ready Visual Exports}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

`agriinsight` includes two innovative "plus" features:
1. **AI Diagnostic Reporting (`report_ai_diagnostics`)**: Automatically parses fitted mixed models, spatial gradients, and treatment effect sizes into structured JSON prompts ready for Large Language Models (LLMs) to draft Results & Discussion sections for *Crop Science*, *Nature Plants*, and *TAG*.
2. **BioRender-Style Vector Export (`export_biorender_svg`)**: Exports multi-layer SVG schematics with colorblind-safe **Okabe-Ito** palettes.

---

```{r setup}
library(agriinsight)
library(lme4)

df_gh <- simulate_agri_data("greenhouse_wheat", n = 96)
mod_gh <- lmer(Biomass_g ~ Inoculant + Pad_Distance + Pot_Density + (1 | Table_Block), data = df_gh)
```

# Generating AI Diagnostic Prompts

```{r ai-report}
ai_prompt <- report_ai_diagnostics(mod_gh, journal = "crop_science", language = "es")

# Print structured prompt string
cat(substr(as.character(ai_prompt), 1, 500), "...\n")
```

---

# Exporting High-End Vector SVG

```{r svg-export}
svg_file <- tempfile(fileext = ".svg")
export_biorender_svg(file = svg_file, theme = "academic_light")
cat("Vector SVG written to:", svg_file, "\n")
```
