---
title: "Endophytes, Soil Microbiome & Biocontrol Kinetics"
author: "Paul Lopez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Endophytes, Soil Microbiome & Biocontrol Kinetics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

Evaluating beneficial plant-associated microorganisms (*Trichoderma*, bacterial endophytes, mycorrhizae) involves:
- Quantifying **root colonization** constrained to $[0, 1]$ (Zero-One Inflated Beta).
- Testing **microbial synergy** in multi-strain consortia via the Bliss Independence Criterion.
- Evaluating biocontrol suppression of soil-borne phytopathogens via **AUDPS** (Area Under Disease Progress Stairs).

---

```{r setup}
library(agriinsight)
```

# 1. Dual Inoculation Synergy (Bliss Independence)

```{r bliss}
# Control: 20 g, Inoculant A: 28 g, Inoculant B: 27 g, Dual Consortium: 48 g, Physiological Max: 55 g
syn <- calc_bliss_synergy(y_control = 20.0, y_inoc_a = 28.0, y_inoc_b = 27.0, y_dual = 48.0, y_max = 55.0)
syn
```

---

# 2. Biocontrol Disease Suppression (AUDPS vs AUDPC)

```{r biocontrol}
days <- c(0, 7, 14, 21, 28)
root_rot_severity <- c(0, 5, 18, 45, 75)

efficacy <- get_biocontrol_efficacy(days, root_rot_severity)
efficacy
```
