---
title: "Crop Nutrition & Soil Fertility: Plateau Calibration and Economic Optimum (EOFR)"
author: "Paul Lopez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Crop Nutrition & Soil Fertility: Plateau Calibration and Economic Optimum (EOFR)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

Calibrating soil nutrient test values and fertilizer response curves requires non-linear plateau models that capture diminishing returns and threshold saturation:

1. **Linear Plateau (von Liebig)**: Rigid law of the minimum.
2. **Quadratic Plateau (Bullock & Bullock 1994)**: Smooth $C^1$ quadratic trajectory with $x_c = -b / (2c)$.
3. **Mitscherlich-Bray Model**: Asymptotic saturation.
4. **Economically Optimum Fertilizer Rate (EOFR)**: Maximizing margin over fertilizer cost.

---

```{r setup}
library(agriinsight)
set.seed(2026)
```

# Plateau Model Calibration

```{r plateau, fig.width = 6.5, fig.height = 4.5}
n_rate <- seq(0, 250, by = 25)
grain_yield <- ifelse(n_rate < 140, 2400 + 36 * n_rate, 2400 + 36 * 140) + rnorm(length(n_rate), 0, 70)

# Fit Linear Plateau
mod_lp <- get_critical_soil_value(n_rate, grain_yield, method = "linear_plateau")
mod_lp

if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot(mod_lp)
}
```

---

# Economically Optimum Fertilizer Rate (EOFR)

```{r eofr}
# Fertilizer price: $1.2/kg N, Grain price: $250/ton ($0.25/kg)
eofr_res <- get_fertilizer_optimum(mod_lp, price_nutrient = 1.2, price_crop = 250)
eofr_res
```
