---
title: "Multi-Environment Trials (MET): AMMI, GGE Biplots, and Stability Metrics"
author: "Paul Lopez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-Environment Trials (MET): AMMI, GGE Biplots, and Stability Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

Evaluating genotype performance across multi-environment trials (MET) requires decomposing Genotype-by-Environment ($G \times E$) interaction variance and identifying stable, high-yielding cultivars.

`agriinsight` provides complete implementations for:
1. **AMMI** (Additive Main Effects and Multiplicative Interaction)
2. **GGE** (Genotype + Genotype-by-Environment) biplots
3. Comprehensive stability indices: **ASV** (AMMI Stability Value), **YSI** (Yield-Stability Index), **Wricke's Ecovalence**, **Shukla's Stability Variance**, and **Finlay-Wilkinson Joint Regression**.

---

```{r setup}
library(agriinsight)
set.seed(2026)

df_met <- simulate_agri_data("met_field_trials")
```

# AMMI Model Fitting & Stability Table

```{r ammi-fit}
ammi_res <- gxe_ammi(df_met, genotype = "Genotype", environment = "Environment", yield = "Yield")

# View principal component decomposition
ammi_res$pca_summary

# View comprehensive stability rankings
head(ammi_res$stability[order(ammi_res$stability$YSI), ], 6)
```

---

# GGE Biplot & S3 Visualization

```{r biplot, fig.width = 6.5, fig.height = 5}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  # Plot AMMI Biplot (IPCA1 vs IPCA2) with environment vectors
  plot(ammi_res, type = "biplot")
}
```

---

# 2-Stage MET Weighting Bridge

```{r stage2}
# Fit single-environment models for each trial location
env_subsets <- split(df_met, df_met$Environment)
models_stage1 <- lapply(env_subsets, function(sub_df) {
  lm(Yield ~ Genotype + factor(Rep), data = sub_df)
})

# Construct Stage 2 dataset with inverse-variance weights
stage2_res <- stage1_to_stage2(models_stage1, genotype_term = "Genotype")
head(stage2_res$stage2_data, 6)
```
